Peptide retention time prediction

### 2014-11-25 retention.tex
Added citation to Oleg's last article

These methods can be roughly divided into four classes: index-based
methods, modeling-based techniques, machine learning approaches and
look-up tables-based approaches. The main ideas underlying each
category are described below, while Table \ref{tab:rtmethods} lists a
few of the most popular retention time prediction methods from each
category.

@@

These methods involve keeping a database of previously observed
retention times for a set of peptides, and use these values as a basis
for retention time predictions. In particular, in the past years the
use of retention standards has become increasingly
widespread \cite{olegstd, irt}. Such standards typically consist of a
small number of peptides covering a wide span of hydrophobicities and
with well studied behaviors under LC-MS/MS. These peptides are often
analyzed together with the sample of interest, and by comparing their
observed and expected behaviors one can obtain valuable information
for a variety of applications.

@@

% SSRcalc was proposed including a number of additional terms; then it was extended
% for a few other column
Following these observations, Krokhin et al. developed
SSRCalc \cite{Krokhin2004}, which is currently the most widely used
index-based retention time predictor. In the initial version of the
algorithm a collection of 346 tryptic peptides was used to derive two
sets of retention coefficients, one corresponding to the N-terminal
and one to all the other positions in a peptide sequence. Furthermore,
correction factors for peptide length and overall hydrophobicity were
included. Subsequent versions of the algorithm included terms
accounting for additional properties such as isoelectric point,
nearest-neighbor effects of charged peptides and propensity to form
helical structures \cite{Krokhin2006}. However, a limitation of the
approach is that the algorithm was optimized for predicting retention
times for certain chromatography setups. Thus, SSRCalc's prediction
ability deteriorates for experimental conditions diverging from the
setup on which the algorithm was optimized \cite{Spicer2007}.

>>>>

These methods can be roughly divided into four classes: look-up
tables-based approaches, index-based methods, modeling-based methods
and machine learning techniques. The main ideas underlying each
category are described below, while Table \ref{tab:rtmethods} lists a
few of the most popular retention time prediction methods from each
category.

@@

These methods involve keeping a database of previously observed
retention times for a set of peptides, and subsequently use these
values as a basis for retention time predictions. In particular, in
the past years the use of retention standards has become increasingly
widespread \cite{olegstd, irt}. Such standards typically consist of a
small number of peptides covering a wide span of hydrophobicities and
with well studied behaviors under LC-MS/MS. These peptides are often
analyzed together with the sample of interest, and by comparing their
observed and expected behaviors one can obtain valuable information
for a variety of applications.

@@

% SSRcalc was proposed including a number of additional terms; then it was extended
% for a few other column
Following these observations, Krokhin et al. developed SSRCalc
\cite{Krokhin2004}, which is currently the most widely used
index-based retention time predictor. In the initial version of the
algorithm a collection of 346 tryptic peptides was used to derive two
sets of retention coefficients, one corresponding to the N-terminal
and one to all the other positions in a peptide sequence. Furthermore,
correction factors for peptide length and overall hydrophobicity were
included. Subsequent versions of the algorithm included terms
accounting for additional properties such as isoelectric point,
nearest-neighbor effects of charged peptides and propensity to form
helical structures \cite{Krokhin2006}. Recently, it was showed that
interactions involving helical structures are highly complex, and thus
more sophisticated models are needed to account for the effects that
such interactions have on peptide retention time \cite{ah}.

A limitation of SSRCacls is that the algorithm was
optimized for predicting retention times for certain chromatography
setups. Thus, SSRCalc's prediction ability deteriorates for
experimental conditions diverging from the setup on which the
algorithm was optimized \cite{Spicer2007}.
